---
title: "Daily Exercise 9/10"
format: html
self-contained: true
execute: 
  echo: true
---
1,2

```{r}
library(tidyverse)
library(visdat)
library(broom)
?airquality
vis_dat(airquality)

airquality <- airquality |>
  distinct()

cleaned_data <- airquality |>
  drop_na()
```
3.
```{r}
model <- lm(Ozone ~ Temp, data = airquality)
  lm(formula = Ozone ~ Temp, data = airquality)
```
I chose temp as the predictor value because I made an estimated guess that it would have a relationship to ozone. In addtion, the temp data did not have any NA values so I would not have to worry about cleaning the data before working with it.

4.
```{r}
summary(model)
```
This does seem like a valid model, as the p-value is small.

5. The R squared is 0.48, which means that 48% of the variability is explained by the regression model.

6.
```{r}
 broom::augment(model) 
```
7.
```{r}
 ggplot(data = airquality,
        aes(x = Temp, y = Ozone)) +
   geom_point() +
   geom_abline(intercept = 0, slope = 1, color = "red") +
   labs(title = "Actual vs. Predicted Ozone Related to Temperature",
        x = "Temperature",
        subtitle = "Correlation between actual and predicted ozone: 0.7") +
   theme_bw()
   
a <- augment(model, newdata = cleaned_data)
 head(data)
 
 paste("Correlation:", round(cor(a$Ozone, a$.fitted),2))
```

